Bipartite Fidelity and Loschmidt Echo of Bosonic Conformal Interface

### 2017-05-01 gnd_dn_lambda.tex
1. Hamiltonian changed to A B notation 2. bmatrix, pmatrix used 3. \tilde{S} changed to \mathcal{S} 4. labeling changed

\begin{figure}[h]
\centering
\includegraphics{fig_gnd_dn_lambda.pdf}
\caption{Partition function of Hamiltonian with DN and $\lambda$ boundary conditions. The width of the strip is $\pi$ due to folding. We unfold the cylinder and the new stripe have $N$ and $D$ boundary conditions on the left and right plus a $\lambda$ junction in the middle. }
\label{Fig in gnd_dn_lambda}
\end{figure}

In this appendix, we would like to calculate the amplitude for the setup shown in Fig.~\ref{Fig in gnd_dn_lambda}. In particular, the unfolded configuration has D/N boundary conditions at $x = \pm \frac{L}{2}$ and linking boundary condition at $x = 0$. The general solutions can be written as
\begin{equation}
\label{Normalized f in gnd_dn_lambda}
f(k, x) = 
\left\lbrace
\begin{aligned}
  A_1 e^{i kt} \cos(kx +\frac{1}{2}kL ) &  \quad x < 0  \\
  A_2 e^{ikt}  \sin(kx - \frac{1}{2}kL ) & \quad x > 0   \\
\end{aligned} \right. 
\end{equation}
As demonstrated in the {\bf\color{red}main text}, if we denote $f(k,x<0)\equiv\phi_1$ and $f(k,x>0)\equiv\phi_2$, the boundary condition at the junction reads
\begin{eqnarray}\begin{aligned}
\frac{\partial_x \phi_1}{ \partial_t \phi_1} = \lambda^2 \frac{\partial_x \phi_2}{ \partial_t \phi_2} = \tan^2 \theta\frac{\partial_x \phi_2}{ \partial_t \phi_2} \quad \theta \in [0,\frac{\pi}{2} ]  
\end{aligned}\end{eqnarray}
which implies
\begin{equation}
\label{Momentum in gnd_dn_lambda}
k = \frac{2\pi}{L}( n \pm \frac{\theta}{\pi} )  \quad n\in{\bf Z}
\end{equation}
It is evident that the momentum $k$ is shifted from integer multiple of $2\pi/L$ due to the $\lambda$ boundary condition in the middle. 
\begin{comment}
Thus, the normalized eigenfunctions are
\begin{equation}
f_n(x) = \sqrt{\frac{2}{L}}
\left\lbrace
\begin{aligned}
  \cos(kx +\frac{1}{2}kL ) &  \quad x < 0  \\
  \pm \sin(kx - \frac{1}{2}kL ) & \quad x > 0   \\
\end{aligned} \right. 
\qquad 
k = \frac{2\pi}{L}( n \pm  \frac{\theta}{\pi} )  \quad n \in \mathbb{Z} 
\end{equation}
Expand the field $\phi = \sum_n \phi_n f_n(x) $, the action and Hamiltonian becomes
\begin{equation}
  S = \frac{g}{2} \int dt \, \sum_{n \in \mathbb{Z} }\left(  \dot{\phi}^2_n + k^2 \phi_n^2 \right) \implies\quad   g \dot{\phi}_n  = \pi_n \quad \implies H =  
\frac{1}{2g}\sum_{n \in \mathbb{Z} } \pi_n^2 + ( kg )^2  \phi_n^2 
\end{equation}
\end{comment}

It is clear that the normalized eigenfunctions in Eq.~\eqref{Normalized f in gnd_dn_lambda} serves as orthonormal basis in mode expansion. Following the same procedure in \onlinecite{di_francesco_conformal_1997}, we have the Hamiltonian as
\begin{equation}
\label{H in gnd_dn_lambda}
H = \frac{1}{2} \sum_{n \in \mathbb{Z} } |k|  (a^{\dagger}_n a_n + \frac{1}{2} )
\end{equation}
where the momentum $k$ is defined in Eq.~\eqref{Momentum in gnd_dn_lambda}, and the creation and annihilation operators defined as usual
\begin{equation}
\begin{aligned}
a_n = \frac{1}{\sqrt{2}} ( \sqrt{ |k|g} \phi_n + \frac{i }{\sqrt{|k|g} }\pi_n  ) \\
a^{\dagger}_n = \frac{1}{\sqrt{2}} ( \sqrt{ |k|g} \phi_n - \frac{i }{\sqrt{|k|g} }\pi_n  ) \\
\end{aligned}
\end{equation}

The Casimir energy is the vacuum energy brought by the finite size of the setup. From Eq.~\eqref{H in gnd_dn_lambda} and define $x\equiv\theta/\pi$ in Eq.~\eqref{Momentum in gnd_dn_lambda}, we have
\begin{equation}
\begin{aligned}
E_c &= \frac{1}{4} \sum_{n \in \mathbb{Z}} | k| = \frac{\pi}{2L} ( \sum_{n \in \mathbb{Z}}  | n + x | + \sum_{n \in \mathbb{Z}}  | n - x |  ) \\  
&= \frac{\pi}{L} \Big[\sum_{n \ge 0 } ( n + x )^{-s} + \sum_{n \ge 0 }  ( n - x)^{-s}  +   x^{-s} \Big]\Big|_{s = -1} \\
&= \frac{\pi}{L} \left[ \zeta_{\rm H}( -1, x ) + \zeta_{\rm H}( -1, x ) +  x \right] \\
&= \frac{1}{2} ( - x^2 + x - \frac{1}{6})
\end{aligned}
\end{equation}
where we used the fact that the unfolded geometry has length $L=2\pi$. Thus the free energy reads
\begin{equation}
F = \beta E_c = - \frac{\beta}{2} B_2( x) 
\end{equation}
which agrees with the boundary state calculation in Eq.{\bf\color{red}where}

>>>>

\begin{figure}[h]
\centering
\includegraphics{fig_gnd_dn_lambda.pdf}
\caption{Partition function of Hamiltonian with DN and $\lambda$ boundary conditions. The width of the strip is $\pi$ due to folding. We unfold the cylinder and the new stripe have $N$ and $D$ boundary conditions on the left and right plus a $\lambda$ junction in the middle. }
\label{fig:Fig_gnd_dn_lambda}
\end{figure}

In this appendix, we would like to calculate the amplitude for the setup shown in Fig.~\ref{fig:Fig_gnd_dn_lambda}. In particular, the unfolded configuration has D/N boundary conditions at $x = \pm \frac{L}{2}$ and linking boundary condition at $x = 0$. The general solutions can be written as
\begin{equation}
\label{eq:normalized_f}
f(k, x) = 
\left\lbrace
\begin{aligned}
  A_1 e^{i kt} \cos\left(kx +\frac{1}{2}kL \right) &  \quad x < 0  \\
  A_2 e^{ikt}  \sin\left(kx - \frac{1}{2}kL \right) & \quad x > 0   \\
\end{aligned} \right. 
\end{equation}
As demonstrated in the {\bf\color{red}main text}, if we denote $f(k,x<0)\equiv\phi_1$ and $f(k,x>0)\equiv\phi_2$, the boundary condition at the junction reads
\begin{eqnarray}\begin{aligned}
\frac{\partial_x \phi_1}{ \partial_t \phi_1} = \lambda^2 \frac{\partial_x \phi_2}{ \partial_t \phi_2} = \tan^2 \theta\frac{\partial_x \phi_2}{ \partial_t \phi_2} \quad \theta \in \left[0,\frac{\pi}{2} \right]  
\end{aligned}\end{eqnarray}
which implies
\begin{equation}
\label{eq:momentum_gnd_dn_lambda}
k = \frac{2\pi}{L}\left( n \pm \frac{\theta}{\pi} \right)  \quad n\in\mathbb{Z}
\end{equation}
It is evident that the momentum $k$ is shifted from integer multiple of $2\pi/L$ due to the $\lambda$ boundary condition in the middle. 
\begin{comment}
Thus, the normalized eigenfunctions are
\begin{equation}
f_n(x) = \sqrt{\frac{2}{L}}
\left\lbrace
\begin{aligned}
  \cos(kx +\frac{1}{2}kL ) &  \quad x < 0  \\
  \pm \sin(kx - \frac{1}{2}kL ) & \quad x > 0   \\
\end{aligned} \right. 
\qquad 
k = \frac{2\pi}{L}\left( n \pm  \frac{\theta}{\pi} \right)  \quad n \in \mathbb{Z} 
\end{equation}
Expand the field $\phi = \sum_n \phi_n f_n(x) $, the action and Hamiltonian becomes
\begin{equation}
  S = \frac{g}{2} \int dt \, \sum_{n \in \mathbb{Z} }\left(  \dot{\phi}^2_n + k^2 \phi_n^2 \right) \implies\quad   g \dot{\phi}_n  = \pi_n \quad \implies H =  
\frac{1}{2g}\sum_{n \in \mathbb{Z} } \pi_n^2 + ( kg )^2  \phi_n^2 
\end{equation}
\end{comment}

It is clear that the normalized eigenfunctions in Eq.~\eqref{eq:normalized_f} serves as orthonormal basis in mode expansion. Following the same procedure in \onlinecite{di_francesco_conformal_1997}, we have the Hamiltonian as
\begin{equation}
\label{eq:H_in_gnd_dn_lambda}
H = \frac{1}{2} \sum_{n \in \mathbb{Z} } |k|  \left(a^{\dagger}_n a_n + \frac{1}{2} \right)
\end{equation}
where the momentum $k$ is defined in Eq.~\eqref{eq:momentum_gnd_dn_lambda}, and the creation and annihilation operators defined as usual
\begin{equation}
\begin{aligned}
a_n = \frac{1}{\sqrt{2}} \left( \sqrt{ |k|g} \phi_n + \frac{i }{\sqrt{|k|g} }\pi_n  \right) \\
a^{\dagger}_n = \frac{1}{\sqrt{2}} \left( \sqrt{ |k|g} \phi_n - \frac{i }{\sqrt{|k|g} }\pi_n  \right) \\
\end{aligned}
\end{equation}

The Casimir energy is the vacuum energy brought by the finite size of the setup. From Eq.~\eqref{eq:H_in_gnd_dn_lambda} and define $x\equiv\theta/\pi$ in Eq.~\eqref{eq:momentum_gnd_dn_lambda}, we have
\begin{equation}
\begin{aligned}
E_c &= \frac{1}{4} \sum_{n \in \mathbb{Z}} | k| = \frac{\pi}{2L} \left( \sum_{n \in \mathbb{Z}}  | n + x | + \sum_{n \in \mathbb{Z}}  | n - x |  \right) \\  
&= \frac{\pi}{L} \Big[\sum_{n \ge 0 } ( n + x )^{-s} + \sum_{n \ge 0 }  ( n - x)^{-s}  +   x^{-s} \Big]\Big|_{s = -1} \\
&= \frac{\pi}{L} \left[ \zeta_{\rm H}( -1, x ) + \zeta_{\rm H}( -1, x ) +  x \right] \\
&= \frac{1}{2} \left( - x^2 + x - \frac{1}{6}\right)
\end{aligned}
\end{equation}
where we used the fact that the unfolded geometry has length $L=2\pi$. Thus the free energy reads
\begin{equation}
F = \beta E_c = - \frac{\beta}{2} B_2( x) 
\end{equation}
which agrees with the boundary state calculation in Eq.{\bf\color{red}where}

### 2017-05-01 pf_of_id.tex
1. Hamiltonian changed to A B notation 2. bmatrix, pmatrix used 3. \tilde{S} changed to \mathcal{S} 4. labeling changed

In this appendix, we provide more details for calculating the amplitude $Z_{ab}$ in Sec.{\bf\color{red}where}. We started to prove the following identity for a real symmetric matrix $M$
\begin{equation}
\label{1st id in app.pf_of_id}
\exp\Big\{- \vec{b}^{\dagger} M \vec{b}  \Big\} \exp \Big\{ \vec{b}^{\dagger} R \bar{\vec{b}}^{\dagger}  \Big\}  = \exp \Big\{ \vec{b}^{\dagger} e^{-M}  R \bar{\vec{b}}^{\dagger}  \Big\} \exp\Big\{- \vec{b}^{\dagger} M \vec{b}  \Big\} 
\end{equation}
where ${\bf b}$ and $\bar{\bf b}$ are vectors of bosonic operators. As emphasized in the {\bf\color{red}main text}, the matrix notation, for example, $\vec{b}^{\dagger} R \bar{\vec{b}}^{\dagger}$ should really mean $\sum_{ij}b^\dagger_iR_{ij}\bar{b}_j^\dagger$ where the dagger does not transpose the vector.

To prove Eq.~\eqref{1st id in app.pf_of_id}, we first consider the special case where $R={\bf I}$. We diagonalize $M = O^{T} \Lambda O $ and rotate the two sets of boson operators to the diagonal basis
\begin{equation}
  \vec{b}^{\dagger}  M \vec{b} = \vec{d}^{\dagger} \Lambda \vec{d}  \quad \vec{d} = O \vec{b} \quad \vec{\bar{d}}^\dagger = O^T \vec{\bar{b}}^\dagger
\end{equation}
where we understood $\vec{\bar{b}}^\dagger$ is a column vector and independent to $\vec{{b}}^\dagger$. Thus the whole expression can be written as
\begin{equation}
\begin{aligned}
  \exp&\Big\{- \vec{b}^{\dagger} M \vec{b}  \Big\} \exp \Big\{ \vec{b}^{\dagger} \bar{\vec{b}}^\dagger  \Big\}  =  
  \exp\Big\{- \vec{d}^{\dagger} \Lambda \vec{d}  \Big\} \exp \Big\{   \vec{d}^{\dagger} \bar{\vec{d}}^\dagger  \Big\} \\
& = \prod_i  \exp\Big\{- \lambda_i d_i^{\dagger} d_i  \Big\} \exp \Big\{  d_i^{\dagger} \bar{d}_i ^{\dagger}  \Big\}
\end{aligned}
\end{equation}
We recall for $ [X, Y] = sY $, 
\begin{equation}
  e^X e^{Y} = e^{\exp (s ) Y} e^{X}
\end{equation}
which is a solvable case of Baker-Campbell-Hausdorff formula. Upon taking $X = -\lambda_i d_i^{\dagger} d_i$, $Y = d_i^{\dagger} \bar{d}^{\dagger}_i$, we have
\begin{equation}
\label{lambda commutator}
[- \lambda_i d_i^{\dagger} d_i, d_i ^{\dagger} \bar{d}_i^{\dagger}] =  - \lambda_i  d_i ^{\dagger} \bar{d}_i^{\dagger} 
\end{equation}
and so $s = - \lambda_i$ for each $\lambda_i$. This enable us to commute those exponentials
\begin{equation}
\begin{aligned}
 \exp\Big\{- \vec{b}^{\dagger} M \vec{b}  \Big\} \exp \Big\{  \vec{b}^{\dagger} \bar{\vec{b}}^\dagger  \Big\}   &= \prod_i \exp \Big\{ e^{- \lambda_i }  d^{\dagger}_i \bar{d}^{\dagger}_i  \Big\}  \exp \Big\{-\lambda_i  d^{\dagger}_i d_i  \Big\} \\
 & = \exp \Big\{ \vec{b}^{\dagger} e^{-M}  \bar{\vec{b}}^\dagger  \Big\} \exp\Big\{- \vec{b}^{\dagger} M \vec{b}  \Big\} 
\end{aligned}
\end{equation}
For the general case where $R \ne I$, we take $\vec{\bar{d}}^* = O^T R \vec{\bar{b}}^*$. This will not change the commutation relation of ${\bf d}$, and the role of $\bar{b}$ is decorative in Eq.~\eqref{lambda commutator}. Hence the rest of the proof follows the same way. \hfill$\blacksquare$

A direct consequence of Eq.~\eqref{1st id in app.pf_of_id} is the following
\begin{equation}
\label{2nd id in app.pf_of_id}
Z_{ab} = \langle 0 | \exp\Big\{ \vec{b} R_a^* \vec{\bar{b}}\Big\} \exp\Big\{ - \vec{b}^{\dagger} M  \vec{b} \Big\}   \exp\Big\{  \vec{b}^{\dagger} R_b  \vec{\bar{b}}^{\dagger}\Big\}  |0  \rangle  = \frac{1}{\det( 1- R_a^{\dagger} e^{-M} R_b )} 
\end{equation}
where $|0\rangle$ is the vacuum for ${\bf b}$ and $\bar{\bf b}$. Using the identity we have just proved, 
\begin{equation}
Z_{ab} =   \langle 0 | \exp\Big\{ \vec{b} R_a^* \vec{\bar{b}}\Big\}  \exp \Big\{ \vec{b}^{\dagger} e^{-M}  R_b \bar{\vec{b}}^{\dagger}  \Big\}  |0 \rangle 
\end{equation}
A direct application of the MacMahon master theorem
\begin{equation}
  \langle 0 | \exp \Big\{ \vec{b}_1 X \vec{b}_2 \Big\}  \exp \Big\{ \vec{b}^{\dagger}_1 Y \vec{b}^{\dagger}_2 \Big\}|0  \rangle 
 = \frac{1}{\det(1 - X^T Y )}
\end{equation}
proves Eq.~\eqref{2nd id in app.pf_of_id}. \hfill$\blacksquare$

>>>>

In this appendix, we provide more details for calculating the amplitude $Z_{ab}$ in Sec.{\bf\color{red}where}. We started to prove the following identity for a real symmetric matrix $M$
\begin{equation}
\label{eq:first_id_app_pf_of_id}
\exp\Big\{- \vec{b}^{\dagger} M \vec{b}  \Big\} \exp \Big\{ \vec{b}^{\dagger} R \bar{\vec{b}}^{\dagger}  \Big\}  = \exp \Big\{ \vec{b}^{\dagger} e^{-M}  R \bar{\vec{b}}^{\dagger}  \Big\} \exp\Big\{- \vec{b}^{\dagger} M \vec{b}  \Big\} 
\end{equation}
where ${\bf b}$ and $\bar{\bf b}$ are vectors of bosonic operators. As emphasized in the {\bf\color{red}main text}, the matrix notation, for example, $\vec{b}^{\dagger} R \bar{\vec{b}}^{\dagger}$ should really mean $\sum_{ij}b^\dagger_iR_{ij}\bar{b}_j^\dagger$ where the dagger does not transpose the vector.

To prove Eq.~\eqref{eq:first_id_app_pf_of_id}, we first consider the special case where $R=\mathbb{I}$. We diagonalize $M = O^{T} \Lambda O $ and rotate the two sets of boson operators to the diagonal basis
\begin{equation}
  \vec{b}^{\dagger}  M \vec{b} = \vec{d}^{\dagger} \Lambda \vec{d}  \quad \vec{d} = O \vec{b} \quad \vec{\bar{d}}^\dagger = O^T \vec{\bar{b}}^\dagger
\end{equation}
where we understood $\vec{\bar{b}}^\dagger$ is a column vector and independent to $\vec{{b}}^\dagger$. Thus the whole expression can be written as
\begin{equation}
\begin{aligned}
  \exp&\Big\{- \vec{b}^{\dagger} M \vec{b}  \Big\} \exp \Big\{ \vec{b}^{\dagger} \bar{\vec{b}}^\dagger  \Big\}  =  
  \exp\Big\{- \vec{d}^{\dagger} \Lambda \vec{d}  \Big\} \exp \Big\{   \vec{d}^{\dagger} \bar{\vec{d}}^\dagger  \Big\} \\
& = \prod_i  \exp\Big\{- \lambda_i d_i^{\dagger} d_i  \Big\} \exp \Big\{  d_i^{\dagger} \bar{d}_i ^{\dagger}  \Big\}
\end{aligned}
\end{equation}
We recall for $ [X, Y] = sY $, 
\begin{equation}
  e^X e^{Y} = e^{\exp (s ) Y} e^{X}
\end{equation}
which is a solvable case of Baker-Campbell-Hausdorff formula. Upon taking $X = -\lambda_i d_i^{\dagger} d_i$, $Y = d_i^{\dagger} \bar{d}^{\dagger}_i$, we have
\begin{equation}
\label{eq:lambda_commutator}
[- \lambda_i d_i^{\dagger} d_i, d_i ^{\dagger} \bar{d}_i^{\dagger}] =  - \lambda_i  d_i ^{\dagger} \bar{d}_i^{\dagger} 
\end{equation}
and so $s = - \lambda_i$ for each $\lambda_i$. This enable us to commute those exponentials
\begin{equation}
\begin{aligned}
 \exp\Big\{- \vec{b}^{\dagger} M \vec{b}  \Big\} \exp \Big\{  \vec{b}^{\dagger} \bar{\vec{b}}^\dagger  \Big\}   &= \prod_i \exp \Big\{ e^{- \lambda_i }  d^{\dagger}_i \bar{d}^{\dagger}_i  \Big\}  \exp \Big\{-\lambda_i  d^{\dagger}_i d_i  \Big\} \\
 & = \exp \Big\{ \vec{b}^{\dagger} e^{-M}  \bar{\vec{b}}^\dagger  \Big\} \exp\Big\{- \vec{b}^{\dagger} M \vec{b}  \Big\} 
\end{aligned}
\end{equation}
For the general case where $R \neq\mathbb{I}$, we take $\vec{\bar{d}}^* = O^T R \vec{\bar{b}}^*$. This will not change the commutation relation of ${\bf d}$, and the role of $\bar{b}$ is decorative in Eq.~\eqref{eq:lambda_commutator}. Hence the rest of the proof follows the same way. \hfill$\blacksquare$

A direct consequence of Eq.~\eqref{eq:first_id_app_pf_of_id} is the following
\begin{equation}
\label{eq:second_id_in_app_pf_of_id}
Z_{ab} = \langle 0 | \exp\Big\{ \vec{b} R_a^* \vec{\bar{b}}\Big\} \exp\Big\{ - \vec{b}^{\dagger} M  \vec{b} \Big\}   \exp\Big\{  \vec{b}^{\dagger} R_b  \vec{\bar{b}}^{\dagger}\Big\}  |0  \rangle  = \frac{1}{\det( 1- R_a^{\dagger} e^{-M} R_b )} 
\end{equation}
where $|0\rangle$ is the vacuum for ${\bf b}$ and $\bar{\bf b}$. Using the identity we have just proved, 
\begin{equation}
Z_{ab} =   \langle 0 | \exp\Big\{ \vec{b} R_a^* \vec{\bar{b}}\Big\}  \exp \Big\{ \vec{b}^{\dagger} e^{-M}  R_b \bar{\vec{b}}^{\dagger}  \Big\}  |0 \rangle 
\end{equation}
A direct application of the MacMahon master theorem
\begin{equation}
  \langle 0 | \exp \Big\{ \vec{b}_1 X \vec{b}_2 \Big\}  \exp \Big\{ \vec{b}^{\dagger}_1 Y \vec{b}^{\dagger}_2 \Big\}|0  \rangle 
 = \frac{1}{\det(1 - X^T Y )}
\end{equation}
proves Eq.~\eqref{eq:second_id_in_app_pf_of_id}. \hfill$\blacksquare$
